Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-02-14 papers/wdmerger_I/paper.tex
Revamp discussion of density resets to be consistent with recent CASTRO changes. Also, remove reference to the dt limiting, which we are no longer doing.

A related issue that required a code improvement was that in cases of
large density gradients such as the edge of a star, it is possible to
generate negative densities in zones even with the more strongly
limited PPM. This can occur if a region of large density is moving
away from an ambient zone at relatively large speeds; then the net
density flux in the ambient zones can be large enough to unphysically
drag the density below zero. In practice, this will occur at the
trailing edge of a star that is moving across a grid. In such a
situation, there are two main approaches one could take: either
explicitly introduce a positivity-guaranteeing diffusive flux, or
reset the characteristics of the affected zone. We choose the latter
approach. Even though it is non-conservative, it preserves a
characteristic we value, which is to keep the edge of the stars
relatively sharp, as they physically should be. Since the mass of the
affected zones is typically already fairly low, this should not
seriously affect the energy conservation properties of our
simulation. Similarly, it is also possible in this physical situation
for very large velocities to be generated in the material behind the
star. The momentum transferred to the ambient zone is close to the
momentum of the stellar material, but this is of much higher density
and so the velocity of the ambient material must become very large to
compensate. While this may be physically viable in certain
circumstances, it poses a problem for our code, which uses subcycling
in time (see below). The timestep is fixed at the beginning of the set
of subcycled timesteps, and if in the middle of the cycle a velocity
is generated that is so large that it violates the CFL criterion for
this timestep, we cannot guarantee the stability of the resulting
solution since we cannot alter the timestep. (Nor would we want to,
because in some cases because the velocity becomes spuriously large
and would drag down the global timestep.) Therefore we also insert a
check at the end of every hydrodynamics timestep that resets the
velocity and/or sound-speed (indirectly, using the temperature) of any
zone that violates the CFL criterion given the timestep we chose at
the beginning of the cycle. In practice, we find that this is mainly
necessary only for simulations in the inertial reference frame of the
white dwarf merger, which we do not normally do for the reasons stated
in Section \ref{sec:kepler}.

>>>>

A related issue that required a code improvement is that in cases of
large density gradients such as the edge of a star, it is possible to
generate negative densities in zones even with the more strongly
limited PPM. This can occur if a region of large density is moving
away from an ambient zone at relatively large speeds; then the net
density flux in the ambient zones can be large enough to unphysically
drag the density below zero. In practice, this will occur at the
trailing edge of a star that is moving across a grid. In such a
situation, there are two main approaches one could take: either
explicitly introduce a positivity-guaranteeing diffusive flux, or
reset the characteristics of the affected zone. We choose the latter
approach. Even though it is non-conservative, it preserves a
characteristic we value, which is to keep the edge of the stars
relatively sharp, as they physically should be. Since the mass of the
affected zones is typically already fairly low, this should not
seriously affect the energy conservation properties of our
simulation. Our strategy for a reset is as follows: when the density of 
a zone is below a pre-determined density floor (which is typically 
$10^{-5}\ \text{g cm}^{-3}$ for our stellar simulations), we look
at all adjacent zones and find the zone with the highest density.
If it is above the density floor, then we set the field values 
(density, momentum, energy, and temperature) of the
resetted zone to be exactly equal to the field values of this 
adjacent zone. If no adjacent zone reaches the density floor, then
the zone is set to the density floor, and given a temperature equal 
to the temperature floor for our simulations (which is typically 
$10^{5}\ \text{K}$ for our stellar simulations). We then recompute 
the thermodynamics to be consistent with these values. The 
velocity of the zone is set to zero. This latter approach will only
occur in very rare situations, and is mainly there as a safeguard.
